Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-03-12 Grassmannian_cells.tex
Intro edits

The results of \cite{rupel} giving a combinatorial construction of counting polynomials for quiver Grassmannians of preprojective/preinjective representations of generalized Kronecker quivers $K(n)$ suggest the following.
\begin{conjecture}
  There exists a torus action on $\Gr_\bfe(P_{m+1}^V)$ such that the affine cell corresponding to a compatible pair $(S_H,S_V)$ has dimension...
\end{conjecture}
\sayD{I need to include a conjecture about the dimensions of cells coming from the combinatorial construction of counting/Poincare polynomials}
\sayT{Insert something about Dyck path combinatorics and the correspondence here}
\sayT{maybe we can erase some parts which show up twice in the introduction?}
\subsection*{Acknowledgements}We would like to thank Giovanni Cerulli Irelli, Hans Franzen, Oliver Lorscheid and Markus Reineke for very fruitful discussions related to this project.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}
\label{sec:covering}
\noindent
We refer to \cite{gab} for an introduction to covering theory.
Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:s(\alpha)\to t(\alpha)$.
A $\CC$-representation $X$ of $Q$ consists of a collection of $\CC$-vector spaces $X_i$ for $i\in Q_0$ and a collection of $\CC$-linear maps $X_\alpha:X_{s(\alpha)}\to X_{t(\alpha)}$ for $\alpha\in Q_1$.
Given $\CC$-representations $X$ and $Y$ of $Q$, a morphism $f:X\to Y$ is a collection of $\CC$-linear maps $f_i:X_i\to Y_i$ for $i\in Q_0$ satisfying $f_{t(\alpha)}\circ X_\alpha=Y_\alpha\circ f_{s(\alpha)}$ for each $\alpha\in Q_1$.
We write $\rep Q$ for the hereditary abelian category of finite-dimensional $\C$-representations of $Q$ and we assume in the following that all representations are finite-dimensional.

@@

Our goal is to give a geometric explanation for this by showing that the compatible pairs provide a natural labeling for the cells of $\Gr_\bfe(P_{m+1})$ found in Theorem~\ref{celldec}.
In fact, we will see more: that the cells of quiver Grassmannians $\Gr_\bfe(P_{m+1}^V)$ for truncated preprojectives $P_{m+1}^V$ are also naturally labeled by compatible pairs.
We accomplish this by providing a bijection between the compatible pairs as in Theorem~\ref{th:combinatorial euler} and the successor closed sets of vertices in the 2-quivers $\cQ_{m+1}^{[r]}$ used in Theorem~\ref{thm:2quivercells} to describe the nonempty cells.
\begin{theorem}
  \label{th:compatible cells}
  \sayD{Fix abuse of notation $H_m$.}
  For $m\ge1$ and $V\in\Gr(\cH_m)$ or $V=0$, each quiver Grassmannian $\Gr_\bfe(P_{m+1}^V)$ admits a cell decomposition with cells labeled by compatible pairs in the maximal Dyck path $D_{m+1}^{[r]}$, where $r=\dim V$.
\end{theorem}
\begin{proof}
  The recursive construction of the 2-quivers $\cQ_{m+1}^{[r]}$ provides a natural ordering of the vertices in the underlying quiver $Q_{m+1}^{[r]}$.
  Indeed, when considering the recursive construction of the 2-quiver $Q_{m+1}^{[r]}$ from \eqref{eq:recursive 2-quivers}, we order the component quivers $Q_{m,i}$ and $Q_{m,*}^{[1]}$ naturally according to their indices so that $Q_{m,*}^{[1]}$ comes last.
  This provides a bijection of these vertices with the edges of $D_{m+1}^{[r]}$ whereby vertices covering the vertex 1 (resp.\ vertex 2) of $K(n)$ correspond to horizontal edges (resp.\ vertical edges) of $D_{m+1}^{[r]}$.

@@

The results of \cite{rupel} provide a stronger statement than Theorem~\ref{th:combinatorial euler}.
Indeed, the compatible pairs are shown to compute the counting polynomials of these quiver Grassmannians $\Gr_\bfe(P_{m+1})$ over a finite field (these coincide with their Poincar\'e polynomials in this case).  
We conjecture that the torus action on $\Gr_\bfe(P_{m+1})$ can be chosen to provide a geometric explaination of this result.
\begin{conjecture}
  For $m\ge1$ and $V\in\Gr(H_m)$ or $V=0$, there exists a torus action on $\Gr_\bfe(P_{m+1}^V)$ such that the dimension of the cell labeled by a compatible pair $(S_H,S_V)$ in the maximal Dyck path $D_{m+1}^{[r]}$, where $r=\dim V$, is given by $\overline{\gamma}_{S_H,S_V}=\sum\limits_{e<e'\in E_{m+1}^{[r]}}\overline{\gamma}_\omega(e,e')$ for
  \begin{equation*}
    \overline{\gamma}_\omega(e,e')=\begin{cases}
      -n & \text{if $e\in S_H$ and $e'\in S_V$;}\\ 
      1 & \text{if $e\in S_H$ and $e'\in H_{m+1}^{[r]}\setminus S_H$;}\\ 
      1 & \text{if $e\in V_{m+1}^{[r]}\setminus S_V$ and $e'\in S_V$;}\\ 
      0 & \text{otherwise.}\\ 
    \end{cases}
  \end{equation*}
\end{conjecture}

>>>>

The results of \cite{rupel} giving a combinatorial construction of counting polynomials for quiver Grassmannians of preprojective/preinjective representations of generalized Kronecker quivers $K(n)$ suggest that the dimensions of cells can be computed using this combinatorics (or the equivalent combinatorics of compatible pairs).
This is made precise in Conjecture~\ref{conj:cell dimensions}.
\sayT{maybe we can erase some parts which show up twice in the introduction?}
\subsection*{Acknowledgements}We would like to thank Giovanni Cerulli Irelli, Hans Franzen, Oliver Lorscheid and Markus Reineke for very fruitful discussions related to this project.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiver Covering Theory}
\label{sec:covering}
\noindent
We refer to \cite{gab} for an introduction to covering theory.
Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:s(\alpha)\to t(\alpha)$.
A $\CC$-representation $X$ of $Q$ consists of a collection of $\CC$-vector spaces $X_i$ for $i\in Q_0$ and a collection of $\CC$-linear maps $X_\alpha:X_{s(\alpha)}\to X_{t(\alpha)}$ for $\alpha\in Q_1$.
Given $\CC$-representations $X$ and $Y$ of $Q$, a morphism $f:X\to Y$ is a collection of $\CC$-linear maps $f_i:X_i\to Y_i$ for $i\in Q_0$ satisfying $f_{t(\alpha)}\circ X_\alpha=Y_\alpha\circ f_{s(\alpha)}$ for each $\alpha\in Q_1$.
We write $\rep Q$ for the hereditary abelian category of finite-dimensional $\C$-representations of $Q$ and we assume in the following that all representations are finite-dimensional.

@@

Our goal is to give a geometric explanation for this by showing that the compatible pairs provide a natural labeling for the cells of $\Gr_\bfe(P_{m+1})$ found in Theorem~\ref{celldec}.
In fact, we will see more: that the cells of quiver Grassmannians $\Gr_\bfe(P_{m+1}^V)$ for truncated preprojectives $P_{m+1}^V$ are also naturally labeled by compatible pairs.
We accomplish this by providing a bijection between the compatible pairs as in Theorem~\ref{th:combinatorial euler} and the successor closed sets of vertices in the 2-quivers $\cQ_{m+1}^{[r]}$ used in Theorem~\ref{thm:2quivercells} to describe the nonempty cells.
\begin{theorem}
  \label{th:compatible cells}
  For $m\ge1$ and $V\in\Gr(\cH_m)$ or $V=0$, each quiver Grassmannian $\Gr_\bfe(P_{m+1}^V)$ admits a cell decomposition with cells labeled by compatible pairs in the maximal Dyck path $D_{m+1}^{[r]}$, where $r=\dim V$.
\end{theorem}
\begin{proof}
  The recursive construction of the 2-quivers $\cQ_{m+1}^{[r]}$ provides a natural ordering of the vertices in the underlying quiver $Q_{m+1}^{[r]}$.
  Indeed, when considering the recursive construction of the 2-quiver $Q_{m+1}^{[r]}$ from \eqref{eq:recursive 2-quivers}, we order the component quivers $Q_{m,i}$ and $Q_{m,*}^{[1]}$ naturally according to their indices so that $Q_{m,*}^{[1]}$ comes last.
  This provides a bijection of these vertices with the edges of $D_{m+1}^{[r]}$ whereby vertices covering the vertex 1 (resp.\ vertex 2) of $K(n)$ correspond to horizontal edges (resp.\ vertical edges) of $D_{m+1}^{[r]}$.

@@

The results of \cite{rupel} provide a stronger statement than Theorem~\ref{th:combinatorial euler}.
Indeed, the compatible pairs are shown to compute the counting polynomials of these quiver Grassmannians $\Gr_\bfe(P_{m+1})$ over a finite field (these coincide with their Poincar\'e polynomials in this case).  
We conjecture that the torus action on $\Gr_\bfe(P_{m+1})$ can be chosen to provide a geometric explaination of this result.
\begin{conjecture}
  \label{conj:cell dimensions}
  For $m\ge1$ and $V\in\Gr(H_m)$ or $V=0$, there exists a torus action on $\Gr_\bfe(P_{m+1}^V)$ such that the dimension of the cell labeled by a compatible pair $(S_H,S_V)$ in the maximal Dyck path $D_{m+1}^{[r]}$, where $r=\dim V$, is given by $\overline{\gamma}_{S_H,S_V}=\sum\limits_{e<e'\in E_{m+1}^{[r]}}\overline{\gamma}_\omega(e,e')$ for
  \begin{equation*}
    \overline{\gamma}_\omega(e,e')=\begin{cases}
      -n & \text{if $e\in S_H$ and $e'\in S_V$;}\\ 
      1 & \text{if $e\in S_H$ and $e'\in H_{m+1}^{[r]}\setminus S_H$;}\\ 
      1 & \text{if $e\in V_{m+1}^{[r]}\setminus S_V$ and $e'\in S_V$;}\\ 
      0 & \text{otherwise.}\\ 
    \end{cases}
  \end{equation*}
\end{conjecture}
